Calculating Biological Module Enrichment or Depletion and Visualizing Data on Large-scale Molecular Maps with \pkg{ACSNMineR} and \pkg{RNaviCell} packages

### 2016-08-26 jss/Deveau.tex
Added script to reproduce "mainz"
- Updated mainz_acsn.png
- Updated results from enrichment, added that we display the 10 most
significant out of 125
- Script run on 25-26/08/2016 on R 3.2.2, windows 10
R version 3.2.2 (2015-08-14)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 8 x64 (build 9200)

locale:
[1] LC_COLLATE=French_France.1252  LC_CTYPE=French_France.1252
LC_MONETARY=French_France.1252 LC_NUMERIC=C
LC_TIME=French_France.1252

attached base packages:
[1] parallel  stats     graphics  grDevices utils     datasets  methods
base

other attached packages:
[1] RNaviCel

dim(results)
[1] 7 9
\end{example}

@@

\begin{tabular}{l c c c c}
\toprule
Module & Mod. size & Nb genes & pval & Cor. pval\\
\midrule
CellCycle & 242 & 20 & 1.70e-03 & 0.021 \\
CellCycle:E2F1\_TARGETS & 130 & 14 & 6.26e-04 & 0.009 \\
CellCycle:E2F2\_TARGETS & 35 & 8 & 4.70e-05 & 0.009 \\
CellCycle:E2F3\_TARGETS & 51 & 11 & 3.82e-06 & 0.0002 \\
CellCycle:E2F4\_TARGETS & 100 & 15 & 9.99e-06 & 0.0003 \\
DNA\_repair:SPINDLE\_CHECKPOINT & 28 & 5 & 3.82e-03 & 0.04 \\
EMT\_motility:ECM & 147 & 13 & 5.09e-03 & 0.045 \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[h!]
  \centering
  \caption{MSigDB canonical pathway database enrichment for genes differentially expressed between ER
positive and ER negative samples in breast cancer.  Module : name of the
module. Mod. size: size of the module. Nb genes: number of genes from input
which are found in the module. Cor. pval: corrected p-value.}
\label{tab:table_msigdb}

\begin{tabular}{l c c c}
\toprule
Pathway & Mod. size & Nb genes & Cor. pval\\
\midrule
KEGG\_CELL\_CYCLE & 128 & 76 & 1.98e-07\\
REACTOME\_CELL\_CYCLE\_MITOTIC & 325 & 158 & 2.40e-07\\
REACTOME\_DNA\_REPLICATION & 192 & 98 & 1.33e-05\\
PID\_FOXM1\_PATHWAY & 40 & 29 & 5.05e-05\\
REACTOME\_MITOTIC\_M\_M\_G1\_PHASES & 172 & 87 & 7.38e-05\\
REACTOME\_CELL\_CYCLE & 421 & 181 & 2.97e-04\\
PID\_AURORA\_B\_PATHWAY & 39 & 27 & 3.38e-04\\
REACTOME\_MITOTIC\_G1\_G1\_S\_PHASES & 137 & 70 & 3.83e-04\\ 
PID\_SYNDECAN\_1\_PATHWAY & 46 & 30 & 4.40e-04\\
REACTOME\_S\_PHASE & 109 & 58 & 4.40e-04\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[!ht]
  \caption{Mean expression values for ER negative differentially expressed
  genes in breast cancer visualized as heatmaps on the ACSN master map.} 
  \centering
  \includegraphics[width=0.8\textwidth]{figures/mainz_acsn.pdf}
  \label{fig:mainz}
\end{figure}

>>>>

dim(results)
[1] 8 9
\end{example}

@@

\begin{tabular}{l c c c c}
\toprule
Module & Mod. size & Nb genes & pval & Cor. pval\\
\midrule
Apoptosis:AKT\_MTOR & 79 & 47 & 0.00 & 0.01 \\ 
CellCycle:E2F2\_TARGETS & 35 & 22 & 0.01 & 0.04 \\ 
CellCycle:E2F3\_TARGETS & 51 & 31 & 0.00 & 0.02 \\ 
CellCycle:E2F4\_TARGETS & 100 & 60 & 0.00 & 0.00 \\ 
DNA\_repair & 346 & 172 & 0.00 & 0.01 \\ 
DNA\_repair:CELL\_CYCLE & 82 & 49 & 0.00 & 0.01 \\ 
DNA\_repair:G1\_CC\_PHASE & 25 & 18 & 0.00 & 0.02 \\ 
DNA\_repair:S\_CC\_PHASE & 46 & 28 & 0.00 & 0.03 \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[h!]
  \centering
  \caption{MSigDB canonical pathway database enrichment for genes differentially expressed between ER
positive and ER negative samples in breast cancer. This table presents the 10 modules with lowest p-value out of 125 with corrected p-value lower than 0.05. Module : name of the
module. Mod. size: size of the module. Nb genes: number of genes from input
which are found in the module. Cor. pval: corrected p-value.}
\label{tab:table_msigdb}

\begin{tabular}{l c c c}
\toprule
Pathway & Mod. size & Nb genes & Cor. pval\\
\midrule
  KEGG\_CELL\_CYCLE & 128 & 76 & $3.9 \times 10^{-8}$ \\ 
  REACTOME\_CELL\_CYCLE\_MITOTIC & 325 & 159 & $3.9 \times 10^{-8}$ \\ 
  REACTOME\_DNA\_REPLICATION & 192 & 98 & $4.9 \times 10^{-6}$ \\ 
  PID\_FOXM1PATHWAY & 40 & 29 & $3.1 \times 10^{-5}$ \\ 
  REACTOME\_MITOTIC\_M\_M\_G1\_PHASES & 172 & 87 & $3.1 \times 10^{-5}$ \\ 
  REACTOME\_CELL\_CYCLE & 421 & 182 & $5 \times 10^{-5}$ \\ 
  REACTOME\_MITOTIC\_G1\_G1\_S\_PHASES & 137 & 71 & $9 \times 10^{-5}$ \\ 
  PID\_AURORA\_B\_PATHWAY & 39 & 27 & 0.0002 \\ 
  REACTOME\_S\_PHASE & 109 & 58 & 0.00024 \\ 
  PID\_SYNDECAN\_1\_PATHWAY & 46 & 30 & 0.00026 \\ 
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[!ht]
  \caption{Mean expression values for ER negative differentially expressed
  genes in breast cancer visualized as heatmaps on the ACSN master map.} 
  \centering
  \includegraphics[width=0.8\textwidth]{figures/mainz_acsn.png}
  \label{fig:mainz}
\end{figure}
